A Computational Account of Default-Mode Function\\ by Control Theory and Reinforcement Learning

### 2016-09-29 paper/paper.tex
more stuff

\section{Reinforcement-Learning as a computational model for DMN function}
\subsection{General Reinforcment-Learning (RL) theory}
In a general reinforcement-learning (RL) framework, an \textit{agent} interacts with an unknown
environment by taking \textit{actions / decisions} $a_t$ which trigger changes in the
\textit{state} of the environment
$s_t \rightarrow s_{t+1}$, accompanied with a real-valued \textit{reward} $r_t$ (or regret!,
since it can be
negative). This should be thought of as an instantaneous satisfaction accompanying the execution of
an action. Here the subscript $t$ denotes time, and can be discrete or continuous.
%% The unknown environment $\mathcal E$ is generally taken as stochastic, but having some kind of
%% memory.
A popular abstract model for the scenario are so-called \textit{Markov Decision Processes (MDPs)}.
This is simply a quintuplet $(\mathcal S, \mathcal A, r, \rho, \mu)$ where
\begin{itemize}
\item $\mathcal S$ is the set of states, e.g $\mathcal S = \{\text{happy}, \text{sad}\}$.
\item $\mathcal A$ is the set of actions, e.g $\mathcal A = \{\text{read}, \text{run},
  \text{laugh}, \text{empathize}, \text{empathize}\}.$
\item $r : \mathcal S \times \mathcal A \rightarrow \mathbb R$, so that $r(s, a)$ is the instant reward for taken action $a$ at state $s$.
\item $\mu: \mathcal S \rightarrow [0, 1]$ is the prior probability on the states so that
  $\mu(s)$ is the probability that the environment starts-off in state $s$.
  \item $\rho : \mathcal S \times \mathcal A \times \mathcal S \rightarrow [0, 1]$, so that
    $\rho(s,a,s') = \mathbb P(s'|s, a)$ is the probability of moving to state $s'$ if action $a$ is taken from state $s$. In addition, one requires that such transitions be Markovian.
\end{itemize}

@@

\textit{Q-learning} (Watkins \& Dayan, 1992), a commonly used off-policy algorithm, uses the
\textit{greedy policy}
\begin{equation}
  \pi(s) = \pi_{\tilde{Q}}(s) = \argmax_{a \in \mathcal A}\tilde{Q}(s, a),
  \label{eq:qlearning}
\end{equation}
where $\tilde{Q}(s, a)$ is the current estimate of the state-action value table.
The hallmark of this trick wherein one only considers deterministic policies is the so-called
\textit{Bellman's equation}
\begin{equation}
  Q^\pi(s_t, a_t) = \mathbb E_{s_{t+1} \sim \rho(.|s_t)} [r(s_t,a_t) + \gamma Q^\pi(s_{t+1}, \pi(s_{t+1}))]
  \label{eq:bellman}
\end{equation}
which is a fixed-point equation for the $Q$-value operator of a deterministic policy $\pi$.
Notice that in the expectation in \eqref{eq:bellman}, the sampling is done over only things which
depend on the environment, and so can be learned off-policy by observing state-transitions
triggered by another behavioral policy, which can be stochastic.

>>>>

\section{Reinforcement-Learning as a computational model for DMN function}
\subsection{General Reinforcment-Learning (RL) theory}
In a general reinforcement-learning (RL) framework, an \textit{agent} interacts with an unknown
environment by taking \textit{actions / decisions} $a_t$ which trigger changes in the
\textit{state} of the environment
$s_t \rightarrow s_{t+1}$, accompanied with a real-valued \textit{reward} $r_t$ (or regret!,
since it can be
negative). This should be thought of as an instantaneous satisfaction accompanying the execution of
an action. Here the subscript $t$ denotes time, and can be discrete or continuous.
%% The unknown environment $\mathcal E$ is generally taken as stochastic, but having some kind of
%% memory.
A popular abstract model for the scenario are so-called \textit{Markov Decision Processes (MDPs)}.
This is simply a quintuplet $(\mathcal S, \mathcal A, r, \rho, \mu)$ where
\begin{itemize}
\item $\mathcal S$ is the set of states, e.g $\mathcal S = \{\text{happy}, \text{sad}\}$.
\item $\mathcal A$ is the set of actions, e.g $\mathcal A = \{\text{read}, \text{run},
  \text{laugh}, \text{empathize}, \text{empathize}\}.$
\item $r : \mathcal S \times \mathcal A \rightarrow \mathbb R$, so that $r(s, a)$ is the instant reward for taking action $a$ at state $s$.
\item $\mu: \mathcal S \rightarrow [0, 1]$ is the prior probability on the states so that
  $\mu(s)$ is the probability that the environment starts-off in state $s$.
  \item $\rho : \mathcal S \times \mathcal A \times \mathcal S \rightarrow [0, 1]$, so that
    $\rho(s,a,s') = \mathbb P(s'|s, a)$ is the probability of moving to state $s'$ if action $a$ is taken from state $s$. In addition, one requires that such transitions be Markovian.
\end{itemize}

@@

\textit{Q-learning} (Watkins \& Dayan, 1992), a commonly used off-policy algorithm, uses the
\textit{greedy policy}
\begin{equation}
  \pi(s) = \pi_{\tilde{Q}}(s) = \argmax_{a \in \mathcal A}\tilde{Q}(s, a),
  \label{eq:qlearning}
\end{equation}
where $\tilde{Q}(s, a)$ is the current estimate of the state-action value table.
The hallmark of this trick wherein one only considers deterministic policies is the so-called
\textit{Bellman's equation}
\begin{equation}
  Q^\pi(s_t, a_t) = \mathbb E_{s_{t+1} \sim \rho(.|s_t)} [r(s_t,a_t) + \gamma Q^\pi(s_{t+1}, \pi(s_{t+1}))]
  \label{eq:bellman}
\end{equation}
which is a fixed-point equation for the deterministic policy $\pi$.
Notice that in the expectation in \eqref{eq:bellman}, the sampling is done over only things which
depend on the environment, and so can be learned off-policy by observing state-transitions
triggered by another behavioral policy, which can be stochastic.
